Kepler's Demographic Cliff: Ages of Young Stars and Planets in the Kepler Field

### 2024-05-02 paper/ms.tex
update ms

Imagine we were to group stars between 0-1\,Gyr, 1-2\,Gyr, and
2-3\,Gyr, and label them ``young'', ``intermediate-age'', and ``old''.
A simple counting exercise from the distributions in
Figure~\ref{fig:hist_tgyro} tells us that there are \ratioobtoybstars\
times as many old stars in the Kepler field as young stars.
Similarly, there are \ratioobtoybplanets\ times as many old planet
hosts as young planet hosts.

In terms of detected planet counts, our rotation-based ages for the
Kepler sample yield \{\nplyounggyro, \nplmidgyro, \nploldgyro\}
detected planets in the 0-1\,Gyr, 1-2\,Gyr, and 2-3\,Gyr bins.  For
the sub-gigayear bin in particular, which holds specific weight in
studies of planet evolution, while there are \nplyounggyro\ planets
with median ages below 1\,Gyr, requiring $t_{\rm gyro}$$<$1\,Gyr at
2$\sigma$ confidence yields \nplyounggyrotwosigma\ planets orbiting
\nplhostsyounggyrotwosigma\ stars.  The stellar and planetary
properties of this latter sample of young planets are given in
Table~\ref{tab:planets}.

@@

@@

these have finite Li ages, but no rotation age (perhaps b/c of the missing prots...)
Kepler-1943 b
 Kepler-101 b
 Kepler-108 b
 Kepler-108 c
  Kepler-63 b
Kepler-1312 b
Kepler-1312 c
   Kepler-1 b

@@

>>>>

Imagine we were to group stars between 0-1\,Gyr, 1-2\,Gyr, and
2-3\,Gyr, and label them ``young'', ``intermediate-age'', and ``old''.
A simple counting exercise from the distributions in
Figure~\ref{fig:hist_tgyro} tells us that there are \ratioobtoybstars\
times as many old stars in the Kepler field as young stars.
Similarly, there are \ratioobtoybplanets\ times as many old planet
hosts as young planet hosts. 
Focusing only on the tails of the distributions (0-0.3\,Gyr and
2.7-3\,Gyr), the implication is that the formation rate of stars in
the Kepler field decreased by a factor of \ratiosfr$\pm$\uncratiosfr\
over the past three billion years.

In terms of detected planet counts, our rotation-based ages for the
Kepler sample yield \{\nplyounggyro, \nplmidgyro, \nploldgyro\}
detected planets in the 0-1\,Gyr, 1-2\,Gyr, and 2-3\,Gyr bins.  For
the sub-gigayear bin in particular, which holds extra weight in
studies of planet evolution, while there are \nplyounggyro\ planets
with median ages below 1\,Gyr, requiring $t_{\rm gyro}$$<$1\,Gyr at
2$\sigma$ confidence yields \nplyounggyrotwosigma\ planets orbiting
\nplhostsyounggyrotwosigma\ stars.  The stellar and planetary
properties of this latter sample of young planets are given in
Table~\ref{tab:planets}.

@@

The fact that only $\approx$90\% of the rotation-based and
lithium-based ages agree

While hopefully clear from our extensive use of quality flags,

@@

these have finite Li ages, but no rotation age (perhaps b/c of the missing prots...)

Kepler-1943 b

 Kepler-101 b

\paragraph{Kepler-108 b and Kepler-108 c}---This sytem of two mutually
inclined giant planets \citep{2017AJ....153...45M} has a host star
that is entering the red giant branch.  This has been previously noted
through asteroseismology \citep{2013ApJ...767..127H}, and an analysis
of the Keck/HIRES spectrum with \texttt{SpecMatch-Synth} concurs, with
$T_{\rm eff}$=5668$\pm$100\,K, $\log g$=3.8$\pm$0.1,
[Fe/H]=0.38$\pm$0.06, and $v\sin i$=3.2$\pm$1.0\kms.

  Kepler-63 b

Kepler-1312 b Kepler-1312 c

   Kepler-1 b

@@

Kepler~108 was reported by \citet{2020ApJ...903...55P} to be a member
of NGC6811 based ``on APOGEE radial velocities and Gaia-DR2 proper
motions''.  The cluster has $\mu_\alpha\approx-3.4$mas/yr and
$\mu_\delta \approx -8.8$mas/yr.  Kepler~108 has
$\mu_\alpha=-1.5\pm0.1$mas/yr and $\mu_\delta \approx
-1.2\pm0.1$mas/yr \citep{Gaia_DR3_2022}.  This is quite discrepant.
